% $Header: /projects/VU-SAGA/Papers/saga_engine_2006/saga_engine.tex,v 1.15 2006/08/13 18:01:00 hkaiser Exp $

\documentclass{acm_proc_article-sp}

\usepackage{ifpdf}
\usepackage{srcltx}
\usepackage{fancyvrb}

\ifpdf
  \usepackage{graphicx}
  \usepackage[pdftex]{hyperref}
  \DeclareGraphicsExtensions{.pdf, .png, .jpg, .JPG}
\else
  \usepackage{graphicx}
  \usepackage[hypertex]{hyperref}
  \DeclareGraphicsExtensions{.ps, .eps}
\fi
  
\newcommand{\I}{\textit}
\newcommand{\B}{\textbf}
\newcommand{\T}{\texttt}

\newcommand{\IB}[1]{\textbf{\textit{#1}}}

\newcommand{\F}[1]{\B{FIXME: #1}}

\newcommand{\up}{\vspace*{-1em}}
\newcommand{\down}{\vspace*{1em}}

\newenvironment{shortlist}{
  \begin{itemize}
  \setlength{\itemsep}{-0.2em}
}{
  \end{itemize}
}

\setlength{\floatsep}{1em}
\setlength{\textfloatsep}{1em}
\setlength{\intextsep}{1.3em}

\DefineShortVerb{\|}
\DefineVerbatimEnvironment{mycode}{Verbatim}
{
  label=Code Example,
  fontsize=\small,
  frame=single,
% framerule=1pt,
  framesep=1em,
% numbers=left,
  gobble=2
}


\pagestyle{plain}

% -------------------------------------------------------------------

\begin{document}

\title{The SAGA C++ Reference Implementation}
\subtitle{Lessons Learnt from Juggling with Seemingly Contradictory Goals}

\numberofauthors{4}
\author{
  \alignauthor Hartmut Kaiser\\
  \affaddr{Center for Computation \& Technology}\\
  \affaddr{Louisiana State University}\\
  \affaddr{Baton Rouge, Louisiana, USA}\\
  \email{\normalsize\T{hkaiser@cct.lsu.edu}}
  \alignauthor Andre Merzky\\
  \affaddr{Vrije Universiteit, Amsterdam}\\
  \affaddr{Amsterdam, The Netherlands}\\
  \email{\normalsize\T{andre@merzky.net}}
  \alignauthor Stephan Hirmer\\
  \affaddr{Center for Computation \& Technology}\\
  \affaddr{Louisiana State University}\\
  \affaddr{Baton Rouge, Louisiana, USA}\\
  \email{\normalsize\T{shirmer@cct.lsu.edu}}
}
\additionalauthors{
  \B{Gabriele Allen}\\
  \hspace*{2em}Center for Computation \& Technology\\
  \hspace*{2em}Louisiana State University\\
  \hspace*{2em}Baton Rouge, Louisiana, USA\\
  \hspace*{2em}email: \texttt{gallen@cct.lsu.edu}
  }

\date{}

\maketitle

\begin{abstract} 

  The Simple API for Grid Applications (SAGA)~\cite{saga_spec} is an ongoing API
  standardization effort within the Open Grid Forum (OGF)~\cite{ogf_website}.  
  The OGF strives to standardize grid middleware, meaning grid service
  interfaces, grid enabled protocols, and grid architecture in
  general.  As of now, most grid standard specifications are in flux,
  and furthermore there exist multiple, incompatible grid middleware
  systems which are deployed in large research or production
  environments.  SAGA is supposed to provide a simple API to
  programmers of scientific applications, allowing them to use higher
  level grid computing paradigms, and thus shielding the programmer
  and the application from the diversity and dynamicity of Grid
  environments.
  
  The SAGA specification itself is expected to extend in scope over
  the next couple of years, in sync with the maturing service
  specifications.  Also, SAGA is defined in SIDL (a language
  independent interface description language~\cite{sidl}), and language bindings for 
  FORTRAN, Java, Python and C are planned, but do not yet exist.

  These 'dynamic' (one could say chaotic) boundary conditions
  make an implementation of the SAGA API specification ... interesting.
  Nevertheless, the perceived need of the grid community for a high
  level API is great enough to tackle that problem \I{now}, and not to
  wait until the standardization landscape settles.

  This paper describes how the C++ SAGA reference implementation tries
  to cope with these boundary conditions -- we think there are lessons
  to learn for other API implementations.  
    
\end{abstract}

% -------------------------------------------------------------------

% \begin{verbatim}
%   - generic API (sync, async, task)
%   - layered architecture (pimpl/facade pattern): 
%     lightweight API objects: API/CPI
%   - sync/async dispatching, asynchronous adaptors
%   - futures
%   - monitoring
%   - trial and error handling
%   - dynamic loading (runtime extensibility): vertical extensibility
%   - static linking possible
%   - packages: horizotal extensibility
%   - macro API definition, adaptor registration
%   - simple but not restricting CPI
%   - adaptor data concept (synchronization of data access)
%   - object persistance
%   - Boost, C++ Std library, portability: cross platform
% \end{verbatim}

  \section{Introduction}
    \label{sec:intro}
    \input{intro.tex}

  \section{Requirements}
    \label{sec:requirements}
    \input{requirements.tex}

  \section{General Design}
    \label{sec:generaldesign}
    \input{generaldesign.tex}

  \section{Implementation Challenges\\ and Details}
    \label{sec:details}
    \input{details.tex}

  \section{Lessons Learnt --\\ Implementation Properties}
    \label{sec:props}
    \input{properties.tex}

	\section{Future Work}
		\label{sec:futurework}
		
	 As repeatedly said, the work on appropriate middleware adaptors will
	 undoubtedly require significant resources in the future -- but
	 without those, the SAGA API will not be usable in real grid
	 environments.  We will work on simplifying adaptor creation,
	 integration and maintenance, and hope for support from the OpenSource
	 community, and from grid middleware vendors.  We deem adaptor
	 development and support more important than the API development
	 itself.
	
	 In parallel, we will strive to provide other language bindings to our
	 implementation, as motivated in section~\ref{ssec:lang}, and want also
	 allow adaptors in various programming languages.  Finally, we plan to
	 apply further generic latency hiding techniques, and to experiment
	 with other API implementations in our framework.

  \section{Conclusion}
    \label{sec:Conclusion}
    \input{conclusion.tex}

% -------------------------------------------------------------------

  \section{Acknowledgments}
  \label{sec:ack}
  
  We thank Thorsten Sch\"utt for his work on Futures; Michel Zsandstra
  and Vladimir Prus for their initial work on dynamic loading; the SAGA 
  groups in OGF for their work on the SAGA specification; the
  Lousisiana State University, the Vrije Universiteit Amsterdam, and
  the XtreemOS project for funding our work.
  % I have to mention XtreemOS from now on...

% -------------------------------------------------------------------

\bibliographystyle{unsrt}
\bibliography{saga_engine}

\end{document}

