\begin{thebibliography}{10}

\bibitem{Pawar}
Bruce~Stubbs Rajesh J.~Pawar, Dongxiao~Zhang and Henry~R. Westrich.
\newblock Preliminary geologic modeling and flow simulation study of co2
  sequestration in a depleted oil reservoir.
\newblock {\em NETL 2001 Conference}, 2001.

\bibitem{Pruess2004}
K.~Pruess, J.~Garc{\'{i}}a, T.~Kovscek, C.~Oldenburg, J.~Rutqvist, C.~Steefel,
  and T.~Xu.
\newblock Code intercomparison builds confidence in numerical simulation models
  for geologic disposal of co2.
\newblock {\em Energy}, 29:1431--1444, 2004.

\bibitem{KalmanPaper}
R.E. Kalman.
\newblock A new approach to linear filteringand prediction problems.
\newblock 2005.

\bibitem{Burger98}
Petr Jan Van~Leeuwen Gerrit~Burgers and Geir Evensen.
\newblock Analysis scheme in the ensemble {K}alman filter.
\newblock {\em Monthly Weather Review, American Meteorological Society}, 126,
  1998.

\bibitem{Zhang}
Yanfen Zhang and Dean Oliver.
\newblock History matching using a hierarchical stochastic model with the
  ensemble {K}alman filter: a field case study.
\newblock {\em SPE Reservoir Simulation Symposium}, 2009.

\bibitem{White1987}
C.~D. White and R.~N. Horne.
\newblock Computing absolute transmissibility in the presence of fine-scale
  heterogeneity.
\newblock In {\em Ninth SPE Symposium on Reservoir Simulation}, pages 209--220,
  San Antonio, Texas, February 1-4 1987.

\bibitem{Durlofsky1992}
L.~J. Durlofsky.
\newblock Modeling fluid flow through complex reservoir beds.
\newblock {\em {SPE} Formation Evaluation}, 7(4):315--322, 1992.

\bibitem{Ballin1993}
P.~R. Ballin, K.~Aziz, A.~G. Journel, and L.~Zuccolo.
\newblock Quantifying the impact of geological uncertainty on reservoir
  performing forecasts.
\newblock In {\em SPE Symposium on Reservoir Simulation}, number 25238-MS in
  SPE Paper, New Orleans, Louisiana, 8 February--3 March 2003.

\bibitem{Gilks96}
W.~R. Gilks, S.~Richardson, and D.~J. Spiegelhalter.
\newblock {\em {M}arkov {C}hain {M}onte {C}arlo in {P}ractice}.
\newblock Chapman and Hall, 1996.

\bibitem{vanLeeuwen2003}
P.~J. van Leeuwen.
\newblock A variance--minimizing filter for large--scale applications.
\newblock {\em Monthly Weather Review}, 131(9):2071--2084, September 2003.

\bibitem{Mitchell02}
P.L.~Houtekamer Herschel L.~Mitchell and Gerard Pellerin.
\newblock Ensemble size, balance and model-error representation in an ensemble
  {K}alman filter.
\newblock {\em Monthly Weather Review, American Meteorological Society}, 130,
  2002.

\bibitem{Anderson2007}
J.~L. Anderson.
\newblock Exploring the need for localization in ensemble data assimilation
  using a hierarchical ensemble filter.
\newblock {\em Physica D}, 230:99--111, 2007.

\bibitem{Michalak2003}
A.~M. Michalak and P.K. Kitanidis.
\newblock A method for enforcing parameter nonnegativity in {B}ayesian inverse
  problems with an application to contaminant source identification.
\newblock {\em Water Resources Research}, 39(2), 2003.

\bibitem{Cloud1}
S.~Jha H.~Kim, Y.~Khamra and M.~Parashar.
\newblock Exploring application and infrastructure adaptaion on hybrid
  grid-cloud infrastructure.
\newblock {\em ScienceCloud: Workshop on Scientific Cloud Computing, in
  conjunction with ACM HPDC-2010}, 2010.

\bibitem{Cloud2}
S.~Jha H.~Kim, Y.~Khamra and M.~Parashar.
\newblock Autonomic approach to integrated hpc grid and cloud usage.
\newblock {\em IEEE Conference on eScience 2009, Oxford}, 2009.

\end{thebibliography}
