\documentclass[]{article}

% Use utf-8 encoding for foreign characters
\usepackage[utf8]{inputenc}

% Setup for fullpage use
\usepackage{fullpage}

% Uncomment some of the following if you use the features
%
% Running Headers and footers
%\usepackage{fancyheadings}

% Multipart figures
%\usepackage{subfigure}

% More symbols
%\usepackage{amsmath}
%\usepackage{amssymb}
%\usepackage{latexsym}

% Surround parts of graphics with box
\usepackage{boxedminipage}

% Package for including code in the document
\usepackage{listings}

% If you want to generate a toc for each chapter (use with book)
\usepackage{minitoc}

% This is now the recommended way for checking for PDFLaTeX:
\usepackage{ifpdf}

\usepackage{longtable}


\ifpdf
\usepackage[pdftex]{graphicx}
\else
\usepackage{graphicx}
\fi

\title{\textbf{Revisions}\\Reliable Replica Exchange Molecular Dynamics Simulation in the Grid using SAGA CPR 
and Migol}

\author{Andre Luckow, Shantenu Jha, Andre Merzky, Bettina Schnor, Joohyun Kim}


\newcounter{mycounter}
\setcounter{mycounter}{0}
\newcommand{\COUNTER}{\addtocounter{mycounter}{1} \arabic{mycounter}}
\begin{document}

\maketitle


\begin{longtable}{|p{0.05\textwidth}|p{0.43\textwidth}|p{0.5\textwidth}|}
\hline 

\textbf{Nb.} &\textbf{Reviewer Comment} &\textbf{Remark} \\
\hline
\endhead
\COUNTER
&Figure 1 adds nothing to the paper and should be removed. This will leave space for (2) and (3) 
below. 
&Figure 1 removed, figure 2/3 placed below of each other.
\\
\hline
\COUNTER
&The second and third paragraphs of Section 2 in the original paper, now removed, should be 
retained. 
& Merged both paragraphs back into section 2.
\\
\hline
\COUNTER
&The sentence,`'``While SAGA represents a well-defined ... even in the presence of failures'', found 
in Section 3 of the original, now removed, should be retained. 
& Added to introduction.
\\
\hline
\COUNTER&
Care needs to exercised in the use of commas throughout. For example on p9, "While it has 
been shown, that resources ..." the comma should be removed. 
&Updated. We carefully proof-read the paper. 
\\
\hline
\COUNTER
& Care needs to be exercised with regard to spelling throughout. For example: on p1, "is an 
important example of a first-generation Grid applications", this should be ``application''; ``hepatatis'' 
should be ``hepatitis''; on p5, ``must initializes'', should be, ``must initialize''. 
& Updated. We carefully proof-read the paper.
\\
\hline
\COUNTER&
On p2, ``We provide irrefutable evidence ...''. Evidence has been provided only for the cases 
studied. There is insufficient rigour in the analysis to merit the use of ``irrefutable''. Please remove, 
``irrefutable''. 
& Updated. We carefully proof-read the paper.
\\
\hline
\COUNTER
&Extension of section 2 and 6. 
&Unfortunately, we are not able to extend these sections due to the strict space limitation of 10 pages.\\
\hline
\end{longtable} 
\end{document}
