%\documentclass[acmjacm,acmnow]{acmtrans2m}
\documentclass[10pt,letterpaper]{article}

\usepackage{graphicx}
\usepackage{url}
\usepackage{color}

\newcommand{\projectnamefull}{\textit{Programming Abstractions for
    Large-Scale Distributed Applications} }

\newcommand{\upup}{\vspace*{-0.5em}}
\newcommand{\up}{\vspace*{-0.25em}}

\newcommand{\I}[1]{\textit{#1}}
\newcommand{\B}[1]{\textbf{#1}}
\newcommand{\T}[1]{\texttt{#1}}

\newcommand{\BI}[1]{\B{\I{#1}}}

% \ifpdf
%   \DeclareGraphicsExtensions{.pdf, .png, .jpg}
% \else
%   \DeclareGraphicsExtensions{.ps, .eps}
% \fi

\long\def\comment#1{{\bf \textcolor{magenta}{\bf #1}}}
\long\def\ccomment#1{{\bf \textcolor{blue}{\bf #1}}}
\newcommand{\C}{\comment}
\newcommand{\CC}{\ccomment}

\newcommand{\yes}{$\bullet$}

\newif\ifdraft

\ifdraft
 \newcommand{\amnote}[1]{   {\textcolor{magenta} { ***Andre:    #1 }}}
 \newcommand{\hknote}[1]{ {\textcolor{blue}    { ***Hartmut:   #1 }}}
 \newcommand{\jhanote}[1]{  {\textcolor{red}     { ***Shantenu: #1 }}}
\else
 \newcommand{\amnote}[1]{}
 \newcommand{\jhanote}[1]{}
 \newcommand{\hknote}[1]{}
\fi

% \markboth{Shantenu Jha et al}{Abstractions for Large-Scale Distributed
%   Applications and Systems}

\begin{document}

\title{SAGA..} 
% \author{Shantenu Jha$^{1,2,3}$, Murray Cole$^{4}$, Daniel
%   S. Katz$^{1,5}$, \\ Manish Parashar$^{6,7}$, Omer Rana$^{8}$,
%   Jon Weissman$^{9}$ \\[1em]
%   \small $^1$Center for Computation \& Technology,
%   Louisiana State University\\[-0.3em]
%   \small $^2$Department of Computer Science,
%   Louisiana State University\\[-0.3em]
%   \small $^3$e-Science Institute,   University of Edinburgh\\[-0.3em]
%   \small $^4$School of Informatics, University of Edinburgh\\[-0.3em]
%   \small $^5$Department of Electrical and Computer Engineering,
%   Louisiana State University\\[-0.3em]
%   \small $^6$NSF Center for Autonomic Computing, Rutgers University\\[-0.3em]
%   \small $^7$Department of  Electrical and Computer Engineering, Rutgers University\\[-0.3em]
%   \small $^8$Department of Computer Science,
%   Cardiff University\\[-0.3em]
%   \small $^9$Department of Computer Science,
%   University of Minnesota\\[-0.3em]
% }

% \category{J.2}{Computer Applications}{Physical Sciences and Engineering}
% \terms{A.1 INTRODUCTORY AND SURVEY, C.2.4 DISTRIBUTED SYSTEMS, D.1.3 Concurrent Programming}
% \keywords{Distributed Applications, Distributed Systems,
%   CyberInfrastructure, Abstractions, Scientific Applications}
%\tableofcontents

\begin{abstract}
  And this is how we changed the world..

\end{abstract}

\maketitle

\begin{verbatim}

Scope of the CPC Paper

1.  Background: changing landscape of computational/digital
  infrastructure:
  - Traditional  Grids	+ Distributed Systems
  - Emerging Distributing Systems (Clouds, distributed data-store..)
  - High-end machines are becoming heterogenous, but at a different
        level of granularity
    Implication of the heterogenity of monolithic systems implies

  Develop scientific applications for all of the above -- in a
  "Standard"  uniform way. This is SAGA i.e, let us not just
  stay confined to traditional Grids. But we need to be very 
  clear that SAGA is applicable only for certain classes of 
  applications; SAGA != MPI; SAGA != a programming model per se
  and not an execution model (like MPI). SAGA supports
  different PM and ExecutionModels.

  Motivation: 
  Traditional Grid Systems:
    - heterogeneity (of what?)
    - .. 
  
   Emerging Distributed Systems:
    - heterogeneity (system interface, but also semantic difference)

   High-End Motivation:
     Why SAGA on these machines?

   SAGA provides the fundamental abstractions --- functionality,... ,
   that are required to develop applications in a system independent
   fashion
 
  Portability: syntactic, semantic and platform independence.

  SAGA Programming System:  (Engine + Packages) + Adaptors + Development tools for SAGA

  The API is exposed via the Engine (functional:context, session,
  non-functional:others) and Packages (functional and non-functional)


2. Some Distributed SCIENTIFIC Applications .. 
   possible use cases..

3. How these features/requirements are supported?
   Could have overlap with "requirements" 
   Develop "requirements"
   ....

4. All the Gory Details that have been developed to support the points
   in section 3.

   This is where OOPSLA paper comes in handy. Lots of the meat 
   is the same, but presented in a different perspective and style.

5. Applications that utilize these... and how.. for what.. (Broad spectrum)
   .. talk about N applications that we've developed
   .. show the specific features from Section 3 and how they are
   addressed/utilised/or contribute to the application usage
  
What is the contribution of this Paper: OOPSLA Paper is Implementation
paper + MCS Paper is Interface Paper, this is a mege of the two using
a consistent vocabulary along with applications.

\end{verbatim}

\section{Introduction} {\bf SJ}

The process of developing and deploying large-scale distributed
applications presents a critical and challenging agenda for
researchers and developers working at the intersection of computer
science, computational science and a diverse range of application
areas. 

\section{Applications}\label{application}{\bf SJ with HK and AM}

Introduce applications -- motivate SAGA

\section{An Overview of SAGA} {\bf AM}

or {\it SAGA in a Nutshell}

\section{Design of the API} {\bf AM}

\section{SAGA Programming System} {\bf HK}

\subsection{Design Philosophy}

A lot of the contact will be similar to the OOPSLA paper, however, we
will have to consistently i) make connections with ``principles of
distributed computing'', ii) make connections with our ``design
decisions'' and (not just objectives), express the tradeoffs and iii)
challenges..

Also discuss PIMPL..  and explain why (critical) it is required
for these (distributed) systems.. (hide implementation details - extensibility,
lightweight facade objects - well integrated with host language,
easy injection of engines functionality)

\subsection{SAGA Engine}

\begin{itemize}
	\item sync, async, task
	\item asynchronous factories
	\item adaptor selection (adaptor registration, error handling and reporting,
	\item on call by call basis)
	\item tasks, task container
	\item bulk optimization
	\item state management
\end{itemize}

\subsection{SAGA Packages}

What more should be said here that isn't just a simple implementation
of the API packages.. 

Q. What to say?  A: OOPSLA paper has a description of the extensibilty
of ``what'' we have, but the why is mimimal. The aim of this section
should be to elaborate on the ``why'', e.g., Vertical layering and
hortizontal layering..  State the above and explain why?

\begin{itemize}
\item Reuse of packages in the adaptors
\item Replica package for example, uses functionality from other packages..
\item ..
\end{itemize}

Earlier packages were translators between the API and the adaptors Now
it has {\bf fall-back functionality} (in the packages)
e.g. Job-service, implementable completely in terms of SAGA and now
almost no other adaptor has this functionality.  This is also tied
into the adaptor selection mechanism.

{\it Parameter Validation:} Some discussion of why it makes
sense to do this in the packages and not multiple times in adaptors

\subsection{Adaptors}

Explain how the diversity of middleware is supported by the adaptors,
e.g,, what is common between GRAM and GridSAM adaptors?

State management:

Error management: External errors and local adaptors...

Introduce the notion/important of CPI ... (at some stage this is a
todo.)

\subsection{Putting it all together: SAGA at work...}{\bf Ole with SJ,
  HK and AM}

{\it Control Flow:} We should have a control flow diagram (akin to the
two diagrams (4 and 6) in the shortened OOPSLA paper)

{\it Timing Information:} Both for single calls, as well as timing
information with increasing load.  For example, take a single, simple
job. Show the time taken for the different stages -- end-to-end.

{\bf Ole to do perfomance analysis}

{\it Deployment and Configuration} Introduce the notion of static
(saga-lite) versus dynamic loading (vanilla saga).

\section{Applications Redux}{\bf SJ with HK and AM}

\subsection{Closing the Loop}

Show how SAGA helps the problems outlined in Section\ref{application}

\subsection{Discuss Applications Further} {\bf SJ}
 
The Kind of Applications that can be developed..

\section*{Acknowledgements}

%\bibliographystyle{unsrt}
\bibliographystyle{phcpc}
\bibliography{saga_cpc}
\end{document}

