% $Header: /projects/VU-SAGA/Papers/saga_engine_2006/saga_engine.tex,v 1.28 2006/10/11 02:55:39 gallen Exp $

\documentclass{acm_proc_article-sp}

\usepackage{ifpdf}
\usepackage{srcltx}
\usepackage{fancyvrb}

\ifpdf
  \usepackage{graphicx}
  \usepackage[pdftex]{hyperref}
  \DeclareGraphicsExtensions{.pdf, .png, .jpg, .JPG}
\else
  \usepackage{graphicx}
  \usepackage[hypertex]{hyperref}
  \DeclareGraphicsExtensions{.ps, .eps}
\fi
  
\newcommand{\I}{\textit}
\newcommand{\B}{\textbf}
\newcommand{\T}{\texttt}

\newcommand{\IB}[1]{\textbf{\textit{#1}}}

\newcommand{\NT}[1]{\normalsize\T{#1}}

\newcommand{\F}[1]{\B{FIXME: #1}}

\newcommand{\up}{\vspace*{-1em}}
\newcommand{\down}{\vspace*{1em}}

\newenvironment{shortlist}{
	\vspace*{-0.85em}
  \begin{itemize}
  \setlength{\itemsep}{-0.3em}
}{
  \end{itemize}
	\vspace*{-0.6em}
}

\setlength{\floatsep}{1em}
\setlength{\textfloatsep}{1em}
\setlength{\intextsep}{1.3em}

\DefineShortVerb{\|}
\DefineVerbatimEnvironment{mycode}{Verbatim}
{
  label=Code Example,
  fontsize=\small,
  frame=single,
% framerule=1pt,
  framesep=1em,
% numbers=left,
  gobble=2
}


\pagestyle{plain}

% -------------------------------------------------------------------

\begin{document}

\title{The SAGA C++ Reference Implementation}
\subtitle{Lessons Learnt from Juggling with Seemingly Contradictory Goals}

\numberofauthors{1}
\author{
  \alignauthor ~ \hfill Hartmut Kaiser             \hfill Andre Merzky           \hfill Stephan Hirmer               \hfill Gabrielle Allen            \hfill ~ \\
  \affaddr{    ~ \hfill Louisiana State University \hfill ~~~~Vrije Universiteit \hfill ~~Louisiana State University \hfill Louisiana State University \hfill ~}\\
  \affaddr{    ~ \hfill Baton Rouge                \hfill Amsterdam              \hfill Baton Rouge                  \hfill Baton Rouge                \hfill ~}\\
  \affaddr{    ~ \hfill Louisiana, USA             \hfill ~~Netherlands          \hfill Louisiana, USA               \hfill Louisiana, USA             \hfill ~}
  \email{      ~ \hfill \NT{hkaiser@cct.lsu.edu    \hfill andre@merzky.net       \hfill shirmer@cct.lsu.edu          \hfill gallen@cct.lsu.edu         \hfill ~}}
}

\date{}

\maketitle

\begin{abstract} 

  The Simple API for Grid Applications (SAGA) is an API
  standardization effort within the Open Grid Forum (OGF).  OGF
  strives to standardize grid middleware and grid architectures. 
   Many OGF specifications are still in flux, and multiple,
  incompatible grid middleware systems are used in research or
  production environments.  SAGA provides a simple API to programmers
  of scientific applications, with high level grid computing
  paradigms which shield from the diversity and dynamic nature of grid
  environments.
  
  The SAGA specification scope will be extended in the coming years, in
  sync with maturing service specifications.  SAGA is defined in SIDL
  (Scientific IDL). A C++ language binding is under development,
  language bindings for FORTRAN, Java, Python and C are planned.

  Implementing the SAGA API specification is an interesting
  and challenging problem itself, due to the dynamic
  environment presented by current grids.
  Nevertheless, the perceived need of the grid community for a high
  level API is great enough to tackle that problem \I{now}, and not to
  wait until the standardization landscape settles.  This paper
  describes how the C++ SAGA reference implementation tries to cope
  with these requirements -- we believe there are lessons to learn for
  other API implementations.  
    
\end{abstract}

% -------------------------------------------------------------------

  \section{Introduction}
    \label{sec:intro}
    \input{intro.tex}

  \section{Requirements}
    \label{sec:requirements}
    \input{requirements.tex}

  \section{General Design}
    \label{sec:generaldesign}
    \input{generaldesign.tex}

  \section{Implementation Details}
    \label{sec:details}
    \input{details.tex}

  \section{Implementation Properties}
    \label{sec:props}
    \input{properties.tex}

	\section{Future Work}
		\label{sec:futurework}
		
	 As mentioned, work on appropriate middleware adaptors will
	 undoubtedly require significant resources in the future.
	 This motivates us to work on simplifying adaptor creation,
	 integration and maintenance, and seek support and contributions from the OpenSource
	 community, and from grid middleware vendors.  
	 We will develop other language bindings on API and adaptor level, and
	 apply further generic latency hiding techniques.

  \section{Conclusion}
    \label{sec:Conclusion}
    \input{conclusion.tex}

% -------------------------------------------------------------------

  \section{Acknowledgments}
  \label{sec:ack}
  
  We thank Thorsten Sch\"utt for his work on Futures; Michel Zsandstra
  and Vladimir Prus for their initial work on dynamic loading; the SAGA 
  groups in OGF for their work on the SAGA specification; the Center 
  for Computation \& Technology at
  Louisiana State University, the Vrije Universiteit Amsterdam, and
  the XtreemOS project for funding this work.

% -------------------------------------------------------------------

\bibliographystyle{unsrt}
\bibliography{saga_engine}

\end{document}

