\documentclass[conference]{IEEEtran}
%\documentclass[conference,final]{IEEEtran}

\input{head}
\input{include}

\begin{document}
%\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
% \conferenceinfo{ECMLS'11,} {June 8, 2011, San Jose, California, USA.}
% \CopyrightYear{2011}
% \crdata{978-1-4503-0702-4/11/06}
% \clubpenalty=10000
% \widowpenalty = 10000

\title{Pilot-Data: An Abstraction for Managing Distributed Data}

\author{}

\date{}
\maketitle

\begin{abstract} 

Alternate title: ``Pilot-Data: An Abstraction for Distributed Data''




\end{abstract}

\section{Introduction and Overview} 




\section{Pilot-Data Overview}

Talk about concept

\begin{itemize}

\item Similar levels of heterogeneity in the data infrastructure

\begin{itemize}
\item File systems, storage, transport protocols, …
\end{itemize}

\item Support application level capabilities to specify dependencies
  at a logical level rather than specific file level

\begin{itemize}
\item First class support for Affinities (D-C, D-D)
\end{itemize}

\item Typically placement and scheduling of data is decoupled from the compute-tasks

\begin{itemize}
\item Integrated approach to compute and data ?
\end{itemize}

\item Dynamic decision for data

\begin{itemize}
\item Analogous  to late-binding of data
\item Fluctuating resources as a fundamental property of DCI
\end{itemize}

\item Abstraction for other factors and not application specific way
\begin{itemize}
\item Varying data sources, fluctuating data rates, etc
\end{itemize}

\end{itemize}












Reasoning about PilotData

Implementation

Substantiate with description of design decisions

\section{Pilot-API}
Pilot-API as a unifying piece


\section{Experiments}

How do we support affinity?

PD on heterogeneous backends

Distributed systems/data

Similar coordination graph as in the past

"Intelligent" data placements:

- ComputeDataService decides whether data, compute or both have to move

- 3 different cases:

i) move all data everywhere

ii) move data when task has been placed

iii) move data and then place task

Assumption:
How is initial data placed?
- distributed
- central in one place

Second application track:
- Data movement in the context of MD jobs

Which application?
- exclusively bfast?
- bwa? another good application?
- in some cases reference data is decomposable...

Look at the difference in $T_c$ for each application in the three
cases. We might introduce the idea of alpha. We dont' need to
quantify, but can qualitatively show how alpha determines the
importance (as measured by time to solution). 

My hunch is that $T_c$ will be more sensitive for data-intensive
application than other way around.

The idea is that for an application which is



\section{Conclusion and Future Work}

\section*{Acknowledgements}
%\up
\footnotesize \footnotesize{This work is funded by NSF CHE-1125332
  (Cyber-enabled Discovery and Innovation), HPCOPS NSF-OCI 0710874
  award, NSF-ExTENCI (OCI-1007115) and NIH Grant Number P20RR016456
  from the NIH National Center For Research Resources. Important
  funding for SAGA has been provided by the UK EPSRC grant number
  GR/D0766171/1 (via OMII-UK) and the Cybertools project (PI Jha)
  NSF/LEQSF (2007-10)-CyberRII-01, NSF EPSCoR Cooperative Agreement
  No. EPS-1003897 with additional support from the Louisiana Board of
  Regents.  SJ acknowledges the e-Science Institute, Edinburgh for
  supporting the research theme. ``Distributed Programming
  Abstractions'' \& 3DPAS. MS is sponsored by the program of BiG Grid,
  the Dutch e-Science Grid, which is financially supported by the
  Netherlands Organisation for Scientific Research, NWO. SJ
  acknowledges useful related discussions with Jon Weissman
  (Minnesota) and Dan Katz (Chicago). We thank J Kim (CCT) for
  assistance with BFAST.  This work has also been made possible thanks
  to computer resources provided by TeraGrid TRAC award TG-MCB090174
  (Jha) and BiG Grid.  This document was developed with support from
  the US NSF under Grant No. 0910812 to Indiana University for
  ``FutureGrid: An Experimental, High-Performance Grid Test-bed''.}

  
\bibliographystyle{IEEEtran}
\bibliography{pilotjob,saga,saga-related}


\end{document}

