\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{url}
\usepackage{float}
\usepackage{times}
\usepackage{multirow}
\usepackage{listings}
\usepackage{times}
\usepackage{paralist}
\usepackage{epsfig}
\usepackage{subfigure}
\usepackage[hypertex]{hyperref}
\usepackage{subfigure}
\usepackage{color}
\usepackage{xspace}

%\documentclass{rspublic}

\usepackage{ifpdf}

\newcommand{\I}[1]{\textit{#1}}
\newcommand{\B}[1]{\textbf{#1}}
\newcommand{\BI}[1]{\textbf{\textit{#1}}}
\newcommand{\T}[1]{\texttt{#1}}

\newcommand{\sagaspec}{\textit{SAGA}\xspace}
\newcommand{\sagaimpl}{\textit{SAGA}\xspace}

\newcommand{\spec}{\sagaspec}
\newcommand{\impl}{\sagaimpl}

\setlength\topmargin{0in}
\setlength\headheight{0in}
\setlength\headsep{0in}
\setlength\textheight{9.5in}
\setlength\textwidth{6.5in}
\setlength\oddsidemargin{0in}
\setlength\evensidemargin{0in}
\setlength\parindent{0.1in}
\setlength\parskip{0.25em}


\ifpdf
 \DeclareGraphicsExtensions{.pdf, .jpg}
\else
 \DeclareGraphicsExtensions{.eps, .ps}
\fi

\newcommand{\note}[1]{ {\textcolor{red} { ***NOTE: #1 }}}

\newif\ifdraft
\drafttrue

\ifdraft
\newcommand{\amnote}[1]{   {\textcolor{magenta} { ***Andre:    #1 }}}
\newcommand{\jhanote}[1]{  {\textcolor{red}     { ***Shantenu: #1 }}}
\else
\newcommand{\amnote}[1]{}
\newcommand{\jhanote}[1]{}
\fi

\begin{document}

 \title{ \Large \vspace{-3.5em} Mephisto Documentation }
 
 \author{ SAGA-Group}
 \date{05-04-2011}
 \maketitle
 

% \jhanote{Remember in addition to serving as an abstract, this will
%   serve as a summary of what will go to the 3 editors of the journals
%   that we are considering publishing a full paper in. Thus some more
%   information/discussion on what the underlying problem and context
%   will be about.}


% \jhanote{Once we have defined / introduced SAGA, we should probably
%   have 3 subsections -- interface, library and adatptors/backends?}

\section*{Introduction}
Mephisto is a perl script that supports SAGA (Simplified API for Grid Applications) installations. SAGA is an API that provides the basic functionality required to build distributed applications, tools and frameworks so as to be independent of the details of the underlying infrastructure ~\cite{}. More information on SAGA can be found on the website:
\begin{verbatim}
  http://www.saga.cct.lsu.edu/
\end{verbatim}

SAGA installations require to install various dependency packages. Mephisto is a tool that automates the installation of SAGA, installing all its dependencies specified for SAGA and SAGA adaptors too.  The version number of all the packages and the type of packages can be specified by the user at the command line. The description of all the command line options available are described in the next section.

Example of the command line statement is : perl mephisto install --target-dir=/home/user/saga --tmp-dir=/home/user/tmp-saga
Other examples with different command line options are explained in later section. 

\section*{Command line options}
This section discusses all the various options available at users disclosure. By default if no options are specified, mephisto pulls up the packages listed on http://static.saga.cct.lsu.edu/mephisto/repository/latest/INDEX. 

Expanded version of the Command to run mephisto: perl mephisto  install/test   --target-dir=/path/to/folder  --tmp-dir=/path/to/folder    --with-saga-version=        --with-boost-version=     --with-globus-version=    --repository=  --with-adaptors=

install                        -  This can be used to install a working copy of SAGA and all its dependencies specified. 
test                             -  This option not only installs SAGA but also runs the command 'make check' to check the installation. It creates a log file displaying the test results.
target-dir                   -  This option is used to specify the path you need to install SAGA and all packages libraries. By default, it creates a folder at '/tmp/meph_inst' 
tmp-dir                       -  This option is used to specify the path for a temporary folder to download all the packages and create log files. By default, it creates a folder at '/tmp/meph_tmp'
repository                  -  Can be used to choose different repositories to install SAGA from. Two repositories available as of now are 'latest' (default) and 'svn_trunk'. Recommended svn_trunk for the                                most recent SAGA version. 
with-saga-version    - Choose saga versions available.
with-boost-version   - Choose boost versions available  >=1.40.0
with-globus-version - Choose globus version available >=4.0. This option would automatically invoke installations of X509, globus adaptors too. 
with-adaptors            - Mention any adaptors to install that would invoke globus installations along with saga-adaptors too. 

\section*{Example Command line options}

\subsection*{- Basic command}
perl mephisto install --target-dir=/path/to/any/desired directory/ --tmp-dir=/path/to/temp/folder 

\subsection*{- Install SAGA from svn_trunk repository}
perl mephisto install --target-dir=/path/to/any/desired directory/ --tmp-dir=/path/to/temp/folder --repository=svn_trunk

\subsection*{- Install SAGA will all saga, globus, boost versions specified}
perl mephisto install --target-dir=/path/to/any/desired directory/ --tmp-dir=/path/to/temp/folder --with-saga-version=1.5.3 --with-boost-version=1.44.0 --with-globus-version=5.0.2
(or)
perl mephisto install --target-dir=/path/to/any/desired directory/ --tmp-dir=/path/to/temp/folder --with-saga=1.5.3 --with-boost=1.44 --with-globus=5.0.2

\subsection*{- Install SAGA with adaptors}
perl mephisto install --target-dir=/path/to/any/desired directory/ --tmp-dir=/path/to/temp/folder --with-adaptors=X509

\subsection*{- How to test a SAGA Core Pre-Release with a specific Boost version (for developers)}
perl mephisto install --target-dir=/path/to/any/desired directory/ --tmp-dir=/path/to/temp/folder --with-saga=1.5.3-pre

\subsection*{- Install and proceed to test the installations}
perl mephisto test --target-dir=/path/to/any/desired directory/ --tmp-dir=/path/to/temp/folder --with-saga-version=1.5.3 --with-boost-version=1.44.0 --with-globus-version=5.0.2

\section*{Example of a Screen Shot after execution}
\section*{Possible errors} 



% 
% \bibliographystyle{IEEEtran} 
% \bibliography{saga}


\end{document}

